Nengo: A Python tool for building large-scale functional brain models

### 2013-09-26 nengo.tex
Added figures for example networks

@@

While the tasks Spaun performs are diverse,
all tasks use a common set of
functional modules that we map
onto the brain areas hypothesized
to perform those functions.
As a result, Spaun is able to match
experimental data across the spectrum
of neuroscientific data,
from the single cell level
to the behavioral level,
despite experimental data
being a very small part of the model description.
Similarly, while this is made possible
by a theoretical framework,
the NEF enables Spaun to employ
many different algorithms with spiking neurons,
rather than defining a single algorithm
that is replicated many times.

@@

%% However, not all simulators can support all neuron types,
%% and all of the attributes that can be
%% set on an ensemble.
%% This is particularly the case
%% when interacting with special-purpose hardware.
%% In these cases, simulators should
%% raise appropriate exceptions
%% when unsupported ensembles are encountered.

@@

\section{Examples and benchmarks}

@@

\begin{figure}
\begin{center}
  Nengo Lorenz code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Nengo Lorenz code listing}
\end{figure}

@@

\begin{figure}
\begin{center}
  PyNN Lorenz code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   PyNN Lorenz code listing}
\end{figure}

\begin{figure}
\begin{center}
  Lorenz attractor benchmarks
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Lorenz benchmarks}
\end{figure}

@@

\begin{figure}
\begin{center}
  Nengo CConv code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Nengo CConv code listing}
\end{figure}

@@

\begin{figure}
\begin{center}
  PyNN CConv code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   PyNN CConv code listing}
\end{figure}

>>>>

%%%%% Document begins

@@

While the tasks Spaun performs are diverse,
all tasks use a common set of
functional modules that we map
onto the brain areas hypothesized
to perform those functions.
Like data-driven large scale models,
Spaun is able to match experimental data
across the neuroscientific spectrum,
from the single cell to the behavioral level.
However, instead of constructing the model
based on data and hoping behavior emerges,
we construct the model
based on behavior
and see that the data emerges.
This is made possible by a theoretical framework
like the memory-prediction framework \cite{TODO},
but unlike it and other frameworks,
the NEF enables Spaun to employ
many different algorithms
based on the behavior being performed,
rather than defining a single algorithm
and hoping that replicating it
many times will enable these behaviors.

@@

@@

\section{Examples}

\subsection{Communication channel}

Faithfully transmit a signal.

\begin{figure}
\begin{center}
  \includegraphics[width=.9\textwidth]{comm_channel}
  \begin{minipage}{0.58\textwidth}
    \begin{lstlisting}
      import nengo
      from nengo.helpers import white_noise
      model = nengo.Model("Communication Channel")
      model.make_node("Input", white_noise(1, 5))
      model.make_ensemble("A", nengo.LIF(30), 1)
      model.make_ensemble("B", nengo.LIF(30), 1)
      model.connect("Input", "A")
      model.connect("A", "B")
      model.probe("Input")
      model.probe("A", filter=0.01)
      model.probe("A.spikes")
      model.probe("B", filter=0.01)
      model.probe("B.spikes")
      sim = model.simulator()
      sim.run(1)
      A_data = sim.data("A")
      A_spikes = sim.data("A.spikes")
      ...
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.37\textwidth}
    \includegraphics[width=\textwidth]{comm_channel_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:cchannel} Figure \arabic{figure}.}{
   Communcation channel model. (Top) Cartoon depiction of the model.
   (Bottom left) Python code to create and run the model.
   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

\begin{figure}
\begin{center}
  \begin{minipage}{.495\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      # We use Nengo to generate
      #     bias: Bias currents
      #     decoders: Decoding weights
      #     encoders: Encoding weights
      # for populations A and B.
      import pynn.nest as pyNN

      lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
                     'tau_syn_I':100}
      dt = 0.5
      pyNN.setup(timestep=dt, min_delay=dt)
      A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
      B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
      input_signal = 0.5
      inputA = [pyNN.DCSource(amplitude=val) for val
                 in A_bias + input_signal * A_encoders]
      for i, pulse in enumerate(inputA):
          pulse.inject_into(A[i:i+1])
      inputB = [pyNN.DCSource(amplitude=val) for val
                 in B_bias]
      for i, pulse in enumerate(inputB):
          pulse.inject_into(B[i:i+1])
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{.495\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      weights = []
      for i in xrange(30):
          for j in xrange(30):
              weights.append((i, j,
                  np.dot(A_decoder[i], B_encoder[j]), 1.0))
      connection = pyNN.Projection(
          A, B, pyNN.FromListConnector(weights))
      A.record('spikes')
      B.record('spikes')
      pyNN.run(1000)

      # Decoded value of B
      B_output = numpy.zeros(1000)
      for i in xrange(1000):
          spikes = B[i:i+1].getSpikes()[:,1]
          spikes = (spikes / dt).astype('int')
          B_output[spikes] += B_decoder[i]
      # Filter with tau = 200 ms
      decay = np.exp(-dt / 200)
      B_output[0, :] *= (1 - decay)
      for i in xrange(1, 1000):
          B_output[i,:] = decay * B_output[i-1,:]
                           + (1-decay) * B_output[i,:]
    \end{lstlisting}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
   An implementation of the communication channel in PyNN.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \begin{minipage}{0.43\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      import nengo
      tau = 0.1; sigma = 10; beta = 8.0/3; rho = 28
      model = nengo.Model('Lorenz attractor')
      model.make_ensemble('State', nengo.LIF(2000),
                            dimensions=3, radius=60)

      def feedback(x):
          dx0 = -sigma * x[0] + sigma * x[1]
          dx1 = -x[0] * x[2] - x[1]
          dx2 = x[0] * x[1] - beta * (x[2] + rho) - rho

          return [dx0 * tau + x[0],
                  dx1 * tau + x[1],
                  dx2 * tau + x[2]]

      model.connect('State', 'State',
                      function=feedback, filter=tau)
      model.probe('State', filter=tau)
      model.probe('State.spikes')
      sim = model.simulator()
      sim.run(6)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.55\textwidth}
    \includegraphics[width=0.95\textwidth]{lorenz}
    \includegraphics[width=\textwidth]{lorenz_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
   Lorenz attractor model. (Left) Python code to create and run the model.
   (Top right) Cartoon depiction of the model.
   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

@@

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.8\textwidth]{cconv}
  \begin{minipage}{0.53\textwidth}
    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
      import nengo
      from nengo.networks import CircularConvolution
      model = nengo.Model("Circular convolution")
      model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
      model.make_ensemble("A", nengo.LIF(512), 4)
      model.connect("Input A", "A")
      model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
      model.make_ensemble("B", nengo.LIF(512), 4)
      model.connect("Input B", "B")
      cconv = model.add(CircularConvolution("Circular Convolution",
          neurons=nengo.LIF(1032), dimensions=4))
      model.connect("A", cconv.A)
      model.connect("B", cconv.B)
      model.make_ensemble("Result", nengo.LIF(512), 4)
      model.connect(cconv, "Result")
      model.probe("A", filter=0.02)
      model.probe("B", filter=0.02)
      model.probe("Result", filter=0.02)
      sim = model.simulator()
      sim.run(0.5)
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.43\textwidth}
    \includegraphics[width=\textwidth]{cconv_res}
  \end{minipage}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
   Circular convolution model. (Top) Cartoon depiction of the model.
   Note that much of the complexity is contained within the
   circular convolution network; the transform matrices shown
   are automatically generated by the network depending on
   input dimensionality.
   (Bottom left) Python code to create and run the model.

   (Bottom right) Results of the model simulation. Lines depict
   the node's output (Input panel) and the decoded value being represented
   by ensembles (A and B panels). Light gray lines in A and B panels denote
   spikes emitted by neurons in the ensembles. Each row depicts spikes
   for one neuron.}
\end{figure}

@@
